Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-06-27 measurements.tex
Rewrittem multicore results based on new measurements

To review the performance of \ac{NC} in a multicore using the \ac{Raspi},
the authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model,
which features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC}
with a 900 MHz clock. Each core has a 32 KiB L1 data cache and 32 KiB
L1 instruction cache. The cores share a 512 KiB L2 cache. All the
measured results, including the baseline results, were obtained with
NEON-enabled code adapted from the Fifi library \cite{fifi}.
The \ac{Raspi} 2 has a NEON extension instruction set which provides
an 128-bit \ac{SIMD} instructions that speed the computations.
Figures~\ref{enc_dec1024}, \ref{enc_dec128} and
\ref{enc_dec16} show the encoding and decoding goodput in MB per
second for different generation sizes, $g = [1024, 128, 16]$
respectively. The size of each coded packet was fixed to $1536$~bytes
since that is the typical size of an Ethernet frame. The blocked operations
were performed dividing the matrices in squared sub-blocks of
$16,\ 32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in
height and width. The figures show only block sizes of
$16 \times 16$, $32 \times 32$, $64 \times 64$ operands since with bigger
block sizes, the operands do not fit in L1 cache. Several testcases
are considered and detailed.

@@

\subsubsection{Encoding blocked}
The encoding results are obtained using the
method described in Section~\ref{sub:implementation-multicore}. The time
recorded includes the dependencies resolving, creation of the \ac{DAG},
and the taskz scheduling. In practice, it would suffice to calculate and
store this information only once per generation size.

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_1024.pdf}
\caption{Encoding and Decoding performance for $g = 1024$. Block size: $32 \times 32$. \cite{wunderlich2015network}}
\label{enc_dec1024}
\end{figure}

For $g= 1024$, the blocked baseline measurements outperforms the non blocked
variant. This means that making the matrix multiplication algorithm cache
efficient brings an increase in goodput by a factor of 2.83. When using the
algorithm described in Section~\ref{sub:implementation-multicore},
encoding with four cores is on average $3.7\times$ faster than with one core.
Similarly, decoding with four codes is $3.05\times$ faster, on average, than
decoding with a single core. Figure~\ref{enc_dec1024} shows that the
implemented algorithm, by exploiting cache efficiency and only three extra
cores provides a $10\times$ gain compared with traditional non-blocked
algorithms. With $g = 1024$, the matrix inversion becomes more expensive
than at smaller generations sizes. Therefore, the decoding goodput
is 46\% of the encoding goodput.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_128.pdf}
\caption{Encoding and Decoding performance for $g = 128$. Block size: $64 \times 64$ \cite{wunderlich2015network}}
\label{enc_dec128}
\end{figure}

For $g = 128$, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is $2.82\times$ faster than with a single core. Due to the smaller
matrix sizes, the gain when using blocked operations in the baselines is not
that significant when compared with $g = 1024$. For the same reason, the matrix
inversion is less expensive. As a consequence, the decoding goodput is 71\%
of the encoding goodput.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_16.pdf}
\caption{Encoding and Decoding performance for $g = 16$. Block size: $16 \times 16$ \cite{wunderlich2015network}}
\label{enc_dec16}
\end{figure}

When $g = 16$, the gains of blocked operations are negligible compared with
the non-blocked ones. The reason behind this behavior is that all the data
fits in the L1 cache. For the scheduled version, since the problem to solve
is so small, the gain when using four cores is a factor of 1.68 compared
with a single core, and 1.01 compared with two cores. Therefore, there
are no practical benefits in using four cores instead of two.

The differences in goodput, for all generation sizes, between the
blocked baseline and the single threaded scheduled measurements are due the
time spent resolving the dependencies and the scheduling overhead. These
effects are negligible for big generation sizes, while considerable for
small matrices.

@@

	16   & 1.703     & 0.345    & 4.9 \\
\hline
	32   & 6.573     & 0.914    & 7.2 \\
\hline
	64   & 21.341    & 3.479    & 6.1 \\
\hline
	128  & 82.326    & 17.411   & 4.7 \\
\hline
	256  & 336.398   & 106.861  & 3.1 \\
\hline
	512  & 1548.750  & 659.469  & 2.3 \\
\hline
	1024 & 6730.380  & 5166.920 & 1.3 \\
\hline
\end{tabular}
\vspace{0.2cm}
\label{runtimes}
\end{table}

>>>>

To review the performance of \ac{NC} in a multicore architecture using the
\ac{Raspi}, the authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model B, which
features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz clock.
Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The cores
share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with NEON-enabled code adapted from the Fifi library
\cite{fifi}. The \ac{Raspi} 2 has a NEON extension instruction set which
provides an 128-bit \ac{SIMD} instructions that speed the computations.
Figures~\ref{enc_dec1024}, \ref{enc_dec128} and \ref{enc_dec16} show the
encoding and decoding goodput in MB per second for different generation sizes,
$g = [1024, 128, 16]$ respectively. For $g = [128, 16]$, the displayed results are the mean values over 1000 measurements, while for $g=1024$, they are mean values over 100 measurements. The size of each coded packet was fixed to
$1536$~bytes since that is the typical size of an Ethernet frame. The blocked
operations were performed dividing the matrices in squared sub-blocks of $16,\
32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in height and
width. The figures show only block sizes of $16 \times 16$, and $128 \times 128$
operands since with bigger block sizes, the operands do not fit in cache.
Several testcases are considered and detailed.

@@

\subsubsection{Encoding blocked}
The encoding results are obtained using the method described in
Section~\ref{sub:implementation-multicore}. The time recorded includes the
dependencies resolving, creation of the \ac{DAG}, and the task scheduling. In
practice, it would suffice to calculate and store this information only once per
generation size.

@@

\begin{figure}[ht!]
\centering
% \includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_1024.pdf}
\includegraphics[width=0.7\textwidth]{images/GenSize_1024_SymbolSize_1536.eps}
\caption{Encoding and Decoding performance for $g = 1024$. Block size: $128 \times 128$.}
\label{enc_dec1024}
\end{figure}

For $g= 1024$, the blocked baseline measurements outperforms the non blocked
variant. This means that making the matrix multiplication algorithm cache
efficient brings an increase in goodput by a factor of 3.24. When using the
algorithm described in Section~\ref{sub:implementation-multicore}, encoding with
four cores is on average $3.9\times$ faster than with one core. Similarly,
decoding with four codes is $3.9\times$ faster, on average, than decoding with
a single core. Figure~\ref{enc_dec1024} shows that the implemented algorithm, by
exploiting cache efficiency and only three extra cores provides a $13\times$
gain compared with traditional non-blocked algorithms. With $g = 1024$, the
matrix inversion becomes more expensive than at smaller generations sizes.
Therefore, the decoding goodput is 58\% of the encoding goodput.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/GenSize_128_SymbolSize_1536.eps}
\caption{Encoding and Decoding performance for $g = 128$. Block size: $128 \times 128$}
\label{enc_dec128}
\end{figure}

For $g = 128$, the differences between the baselines operations show that a
blocked algorithm is 8\% faster than the non-blocked variant. Encoding with
four cores is $2.89\times$ faster than with a single core. Due to the smaller
matrix sizes, the gain when using blocked operations in the baselines is not
that significant when compared with $g = 1024$. For the same reason, the matrix
inversion is less expensive. As a consequence, the decoding goodput is 46\%
of the encoding goodput.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/GenSize_16_SymbolSize_1536.eps}
\caption{Encoding and Decoding performance for $g = 16$. Block size: $16 \times 16$}
\label{enc_dec16}
\end{figure}

When $g = 16$, the gains of blocked operations are negligible compared with
the non-blocked ones. The reason behind this behavior is that all the data
fits in the L1 cache. For the scheduled version, since the problem to solve
is so small, the gain when using four cores is a factor of 2.45 compared
with a single core, and 1.46 compared with two cores. Therefore, the
practical benefits in using four cores instead of two are reduced.

The differences in goodput, for all generation sizes, between the
blocked baseline and the single threaded scheduled measurements are due the
time spent resolving the dependencies and the scheduling overhead. These
effects are negligible for big generation sizes, while considerable for
small matrices. For instance, Figure~\ref{enc_dec16} shows that the encoding
speed when using one core with the described algorithm is 78\% the encoding
speed without the recording and calculation of the \ac{DAG}.

@@

	16   & 1.495     & 0.169    & 8.8 \\
\hline
	32   & 5.365     & 0.514    & 10.4 \\
\hline
	64   & 20.573    & 2.024    & 10.1 \\
\hline
	128  & 81.357    & 11.755   & 6.9 \\
\hline
	256  & 326.587   & 75.451  & 4.3 \\
\hline
	512  & 1354.012  & 540.469  & 2.5 \\
\hline
	1024 & 5965.284  & 4373.329 & 1.3 \\
\hline
\end{tabular}
\vspace{0.2cm}
\label{runtimes}
\end{table}
